Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
consistent domain titles: "list processing" and "text editing"

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library (DSL) of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which DSL components are
useful for which kinds of tasks.
Together, the DSL and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
%% \begin{figure}
%%   \begin{tabular}{lcl}
%%     \toprule
%%     Domain&Example Task&Part of the learned DSL\\\midrule
%%     Lists&
%%     \begin{tabular}{rcl}
%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
%%     \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
%%           \hspace{0.5cm} \emph{(appends lists)}
%%       \end{tabular}

@@

\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Strings}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other ($f_1$/$f_2$ in Strings \& Regression)}
\end{figure}

@@

\subsection{List Functions}\label{listSection}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 244 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 244 tasks.
Our data set primarily consists of arithmetic operations upon sequences,
and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
\code{3}, \code{4}, \code{5}.

@@

\subsection{String Editing}\label{textSection}
Synthesizing programs that edit strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other systems successful.

@@

We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (string editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.

@@

  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\
  Solve time&  29s&s&80s&--&235s

>>>>

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good text editing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library (DSL) of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which DSL components are
useful for which kinds of tasks.
Together, the DSL and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

We apply \system to four domains:
list processing;
text editing (in the style of FlashFill~\cite{gulwani2011automating});
and symbolic regression.
For each of these we initially provide a generic set of programming primitives.
Our algorithm then discovers its own DSL for expressing solutions in the
domain (Tbl.~\ref{initialExampleDSL}).
%% \begin{figure}
%%   \begin{tabular}{lcl}
%%     \toprule
%%     Domain&Example Task&Part of the learned DSL\\\midrule
%%     Lists&
%%     \begin{tabular}{rcl}
%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
%%     \end{tabular}
%%     &
%%     \begin{tabular}{l}
%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
%%           \hspace{0.5cm} \emph{(appends lists)}
%%       \end{tabular}

@@

\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs
  \system discovers for them. Bottom: Subset of learned DSL. Notice that
  learned DSL primitives can call each other
  ($f_0$/$f_1$ in Lists, $f_1$/$f_2$ in Text \& Regression)}
\end{figure}

@@

\subsection{List Processing}\label{listSection}
Synthesizing programs that manipulate data structures is a widely studied
problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of learning functions that
manipulate lists.
We created 236 Lisp-style list manipulation tasks, each with 15 input/output
examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions: many of the functions are
very complicated, and the agent must learn to solve these complicated
problems from only 236 tasks.
Our data set primarily consists of arithmetic operations upon sequences, and
so, in addition to the Lisp primitives provided for the text editing
experiments, we additionally start the system out with the following
primitives:
\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
\code{3}, \code{4}, \code{5}.

@@

\subsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other systems successful.

@@

We automatically generated 109 text editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some text editing learned subroutines }\label{textPrimitives}
  \end{table}

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (text editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.

@@

  \\\midrule\multicolumn{6}{c}{\emph{List Processing}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s

  \\\midrule\multicolumn{6}{c}{\emph{Text Editing}}\\\midrule
  \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\
  Solve time&  29s&s&80s&--&235s
